High-Dimensional Regularized Discriminant Analysis

### 2016-01-30 reports/paper-hdrda.Rnw
Applied edits from Mrs. Young #23

\cite{Friedman:1989tm} proposed the popular regularized discriminant analysis
(RDA) classifier that utilizes a biased covariance-matrix estimator that
partially pools the sample covariance matrices from linear and quadratic
disriminant analysis and shrinks the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure using a
grid search. We formulate a new RDA-based classifier for the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to a classifier in a subspace having a much lower dimension. As a
result, the utilization of the dimension-reduction step yields a substantial
reduction in computation during model selection. Also, our parameterization
offers interpretability that was previously lacking with the RDA classifier. We
demonstrate that our proposed classifier is often superior to several recently
proposed sparse and regularized classifiers in terms of classification accuracy
with three artificial and six real high-dimensional data sets. Finally, we
provide an implementation of our proposed classifier in the {\tt sparsediscrim}
R package, which is available on CRAN.

@@

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
parameter restrictions. Covariance-matrix regularization methods are often based
on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
\cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
\cite{Ye:2005uu} have proposed regularization methods to improve the estimation
of the class covariance matrices by adjusting the eigenvalues of the sample
covariance matrices to ensure positive definiteness. Alternatively, dimension
reduction typically transforms the original data in the high-dimensional space
to a subspace having a much smaller dimension such that redundant and noisy
information is removed
\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed the features within each class are independent so that the population
covariance matrices are diagonal, resulting in positive-definite sample
covariance matrices. For excellent overviews of the regularization problem in
discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
\cite*{Mkhadri:1997gy}.

@@

Our proposed \emph{HDRDA} classifier employs a biased covariance-matrix
estimator that partially pools the individual sample covariance matrices from
the \emph{QDA} classifier with the pooled sample covariance matrix from the
\emph{LDA} classifier. We then shrink the resulting covariance-matrix estimator
towards a scaled identity matrix to ensure positive definiteness. By
reparameterizing the \emph{RDA} classifier similar to \cite*{Hastie:2008dt} and
\cite{Halbe:2007ea}, we show that the pooling parameter in the \emph{HDRDA}
classifier determines the contribution of each training observation to the
estimation of each class covariance matrix, enabling interpretability that has
been previously lacking with the \emph{RDA} classifier \citep{Bensmail:1996fw}.

Next, we achieve a substantial reduction in dimension that greatly reduces the
computational effort during model selection when $p \gg N$ by applying reasoning
similar to \cite{Ye:2006tq}. We establish that the matrix operations
corresponding to the null space of the pooled sample covariance matrix are
redundant and can be discarded from the \emph{HDRDA} decision rule without loss
of classificatory information. Moreover, because the \emph{HDRDA} decision rule
is invariant to adjustments to the zero eigenvalues, the decision rule in the
original feature space is equivalent to a decision rule in a lower dimension,
such that matrix inverses and determinants of relatively small matrices can be
rapidly computed. Also, we provide an efficient algorithm along with pseudocode
to estimate the \emph{HDRDA} classifier's tuning parameters in a grid search via
cross-validation. Finally, we show that several shrinkage methods that are
special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such methods include work from
\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.

The remainder of this paper is organized as follows. In Section
\ref{sec:preliminaries} we present the discrimination problem and necessary
notation to describe our contributions. In Section \ref{sec:rda} we present the
\emph{HDRDA} classifier along with its interpretation, properties, and
computationally efficient model-selection procedure. In Section \ref{sec:sims},
we describe our simulation studies of artificial and real data sets and also
describes the experimental results. We conclude with a brief discussion in
Section \ref{sec:discussion}.

@@

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and provide reduced variability for
$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix regularization applies
  a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
\emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
{\tt sparsediscrim} R package, which is available on CRAN.

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ as it would incorporate $K$ additional tuning parameters to estimate
via cross-validation, which is counterproductive to our goal of improving the
\emph{RDA} classifier's computational efficiency. By setting $\alpha_k$ to an
appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
estimators proposed in the literature can be obtained by selecting $\gamma$
accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.

Perhaps the most well-known of the ridge-like estimators is from
\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
proposed the biased covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.

Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but yields an
interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
as we discuss below, our formulation of \emph{HDRDA} classifier lends itself to
rapid estimation in high-dimensional classification.

@@

Next, we establish properties of the covariance-matrix estimator and the
decision rule employed in the \emph{HDRDA} classifier. By doing so, we
demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
the first summand consists of matrix operations applied to low-dimensional
matrices and the second summand corresponds to the null space of $\widehat{\bm
  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
forms across all classes and can be omitted. For $p \gg N$, the constant
component involves determinants and inverses of high-dimensional matrices, and
by ignoring these calculations, a substantial reduction in computational costs
is achieved. Furthermore, a byproduct is that adjustments to the associated
eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
value decomposition to efficiently calculate the eigenvalue decomposition of
$\widehat{\bm \Sigma}$, further reducing the computational costs of the
\emph{HDRDA} classifier.

@@

As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
have the following corollary.

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

@@

We now present our main result where we decompose \eqref{eq:hdrda} and show that
the term requiring the largest computational costs does not contribute to the
classification of an unlabeled observation performed using Lemma
\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
equivalent, more computationally efficient decision rule.

@@

As we can see, we are able to avoid the time-consuming inverses and determinants
of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculate
these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
with a substantial reduction in computational complexity. Here, we demonstrate
that the computational efficiency in calculating the inverse and determinant of
$\bm W_k$ can be further improved via standard matrix operations by showing that
the inverses and determinants of $\bm W_k$ can be performed on matrices of size
$n_k \times n_k$.

@@

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe an
efficient model-selection procedure along with pseudocode in Algorithm
\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$ that
attains the minimum error rate. To calculate the $V$-fold cross-validation, we
partition the original training data into $V$ mutually exclusive and exhaustive
folds that have approximately the same number of observations. Then, for $v = 1,
\ldots, V$, we classify the observations in the $v$th fold by training a
classifier on the remaining $V - 1$ folds. We then calculate the
cross-validation error as the proportion of misclassified observations across
the $V$ folds.

@@

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$ without repeating the costly computations. When calculating the
discriminant scores of each test observation in $\bm X_{test}$, rather than
calculating each score individually, we can more efficiently calculate the
quadratic forms via a single matrix operation. That is, rather than calculating
$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
individually for each row $\bm x$ of $\bm X_{test}$, we can instead calculate
$(\bm X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm
X_{test} - \xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix
contain the individual quadratic form of each test observation, $\bm x_t$.

@@

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
\Sigma_k$ given in \eqref{eq:block-sigma}. We set the mean vector of class 1 to
$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
features of $\bm \mu_2$ were 1/2, while the rest were set to 0, i.e., $\bm \mu_2
= (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p -
  100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
populations differed in their mean vectors in the first 100 features
corresponding to the first block. No difference in the means occurred in the
remaining blocks.

>>>>

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
that partially pools the sample covariance matrices from linear and quadratic
disriminant analysis and shrinks the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure that uses
a grid search. We formulate a new RDA-based classifier for the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to a classifier in a subspace having a much lower dimension. As a
result, the utilization of the dimension-reduction step yields a substantial
reduction in computation during model selection. Also, our parameterization
offers interpretability that was previously lacking with the RDA classifier. We
demonstrate that our proposed classifier is often superior to several recently
proposed sparse and regularized classifiers in terms of classification accuracy
with three artificial and six real high-dimensional data sets. Finally, we
provide an implementation of our proposed classifier in the {\tt sparsediscrim}
R package, which is available on CRAN.

@@

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
parameter restrictions. Covariance-matrix regularization methods are often based
on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
Additionally, \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
improve the estimation of the class covariance matrices by adjusting the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. Alternatively, dimension reduction typically transforms the
original data in the high-dimensional space to a subspace having a much smaller
dimension such that redundant and noisy information is removed
\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed that the features within each class are independent so that the
population covariance matrices are diagonal, resulting in positive-definite
sample covariance matrices. For excellent overviews of the regularization
problem in discriminant analysis and other techniques, see \cite{Murphy:2012uq}
and \cite*{Mkhadri:1997gy}.

@@

Our proposed \emph{HDRDA} classifier employs a biased covariance-matrix
estimator that partially pools the individual sample covariance matrices from
the \emph{QDA} classifier with the pooled sample covariance matrix from the
\emph{LDA} classifier. We then shrink the resulting covariance-matrix estimator
towards a scaled identity matrix to ensure positive definiteness. By
reparameterizing the \emph{RDA} classifier similar to that of
\cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we show that the pooling
parameter in the \emph{HDRDA} classifier determines the contribution of each
training observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}.

Next, we achieve a substantial reduction in dimension that greatly reduces the
computational effort during model selection when $p \gg N$ by applying reasoning
similar to that of \cite{Ye:2006tq}. We establish that the matrix operations
corresponding to the null space of the pooled sample covariance matrix are
redundant and can be discarded from the \emph{HDRDA} decision rule without loss
of classificatory information. Moreover, because the \emph{HDRDA} decision rule
is invariant to adjustments to the zero eigenvalues, the decision rule in the
original feature space is equivalent to a decision rule in a lower dimension,
such that matrix inverses and determinants of relatively small matrices can be
rapidly computed. Also, we provide an efficient algorithm along with pseudocode
to estimate the \emph{HDRDA} classifier's tuning parameters in a grid search via
cross-validation. Finally, we show that several shrinkage methods that are
special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such techniques include work
from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.

The remainder of this paper is organized as follows. In Section
\ref{sec:preliminaries} we present the discrimination problem and necessary
notation to describe our contributions. In Section \ref{sec:rda} we present the
\emph{HDRDA} classifier along with its interpretation, properties, and
computationally efficient model-selection procedure. In Section \ref{sec:sims}
we describe our simulation studies of artificial and real data sets and describe
the experimental results. We conclude with a brief discussion in Section
\ref{sec:discussion}.

@@

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and provides reduced variability for
$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
regularization applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

@@

As we have discussed above, several eigenvalue adjustment methods have been
proposed that increase eigenvalues (approximately) equal to 0. To further
improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
inverse, we define the eigenvalue adjustment of
\eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
\emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
{\tt sparsediscrim} R package, which is available on CRAN.

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ because the estimation of $K$ additional tuning parameters via
cross-validation is counterproductive to our goal of improving the \emph{RDA}
classifier's computational efficiency. By setting $\alpha_k$ equal to an
appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
estimators proposed in the literature can be obtained when one selects $\gamma$
accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.

Perhaps the most well-known of the ridge-like estimators is from
\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
proposed the biased covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, \cite{Friedman:1989tm}
established the RDA classifier by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.

Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but yields an
interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
as we discuss below, our formulation of the \emph{HDRDA} classifier lends itself
to rapid estimation in high-dimensional classification.

@@

Next, we establish properties of the covariance-matrix estimator and the
decision rule employed in the \emph{HDRDA} classifier. By doing so, we
demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
the first summand consists of matrix operations applied to low-dimensional
matrices and the second summand corresponds to the null space of $\widehat{\bm
  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
forms across all classes and can be omitted. For $p \gg N$, the constant
component involves determinants and inverses of high-dimensional matrices, and
by ignoring these calculations, we achieve a substantial reduction in
computational costs. Furthermore, a byproduct is that adjustments to the
associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize
the singular value decomposition to efficiently calculate the eigenvalue
decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
costs of the \emph{HDRDA} classifier.

@@

As an immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the
following corollary.

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows when we set $\gamma = 0$ in Lemma
\ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

@@

We now present our main result, where we decompose \eqref{eq:hdrda} and show
that the term requiring the largest computational costs does not contribute to
the classification of an unlabeled observation performed using Lemma
\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
equivalent, more computationally efficient decision rule.

@@

Thus, we can avoid the time-consuming inverses and determinants of $p \times p$
covariance matrices in \eqref{eq:hdrda} and instead calculate these same
operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
with a substantial reduction in computational complexity. Here, we demonstrate
that the computational efficiency in calculating the inverse and determinant of
$\bm W_k$ can be further improved via standard matrix operations when we show
that the inverses and determinants of $\bm W_k$ can be performed on matrices of
size $n_k \times n_k$.

@@

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe an
efficient model-selection procedure along with pseudocode in Algorithm
\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$, which
attains the minimum error rate. To calculate the $V$-fold cross-validation, we
partition the original training data into $V$ mutually exclusive and exhaustive
folds that have approximately the same number of observations. Then, for $v = 1,
\ldots, V$, we classify the observations in the $v$th fold by training a
classifier on the remaining $V - 1$ folds. We calculate the cross-validation
error as the proportion of misclassified observations across the $V$ folds.

@@

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$ without repeating the costly computations. When calculating the
discriminant scores of each test observation in $\bm X_{test}$, rather than
calculating each score individually, we can more efficiently calculate the
quadratic forms via a single matrix operation. That is, rather than calculating
$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
individually for each row $\bm x$ of $\bm X_{test}$, we can calculate $(\bm
X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm X_{test} -
\xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix contain the
individual quadratic form of each test observation, $\bm x_t$.

@@

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
\Sigma_k$ given in \eqref{eq:block-sigma} and set the mean vector of class 1 to
$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
features of $\bm \mu_2$ were set to 1/2, while the rest were set to 0, i.e.,
$\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
populations differed in their mean vectors in the first 100 features
corresponding to the first block. No difference in the means occurred in the
remaining blocks.
